Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-07 arxiv-version/introduction.tex
Some clean ups

We study the problem of \textsc{Online Multiclass Linear
Classification with Bandit Feedback}~\cite{Kakade-Shalev-Shwartz-Tewari-2008}.
The problem can be viewed as a game between the learner and the environment:
at each timestep $t$, the environment chooses an example $(x_t, y_t)$, where its feature
part $x_t \in \R^d$ is revealed to the learner.
\footnote{All our results apply to also to a simpler model of
the so called \emph{oblivious adversary} where the whole sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is chosen before the first
round of the game.}
Upon receiving this example, the
learner is asked to make a prediction $\hat{y}_t$.
the learner receives feedback immediately after making its prediction.
In contrast with the standard full information setting (where the feedback given is the correct label $y_t$), the feedback given is only a binary indicator
whether the prediction was correct or not. The protocol of the problem is formally stated below.

%In this problem, in each round, the
%learner receives a feature vector and is asked to predict one of $K$
%possible classes. To allow
%for meaningful use of randomization, we assume that the correct label in a
%particular round is chosen prior to algorithm's prediction in that
%round.
%Immediately after the learner makes its prediction, it
%receives feedback whether the prediction was correct or not. However, in
%contrast with the standard, so called full information setting, the algorithm
%does \emph{not} receive the correct label. The feedback is only binary.

@@

The performance is of the learner is measured by its cumulative number of mistakes
$\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$. The bandit feedback
 is a natural feedback model motivated by applications including
online advertising and online content optimization, where the learner only gets to
see the loss of its action taken.
%where the binary feedback corresponds to a click
%The
%corresponding problem with full information feedback is solved by the classical
%\textsc{Multiclass Perceptron} algorithm and its variants (see e.g.
%\cite{Crammer-Dekel-Keshet-Shalev-Shwartz-Singer-2006}).

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}.
\chicheng{I wonder if we should introduce the two notions of linear separability here,
because strong linear separability is also worth mentioning in the full-information setting.
}

>>>>

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a game between the learner and the adversary: at each timestep $t$,
the adversary chooses an example $(x_t, y_t)$ where the feature vector $x_t \in
\R^d$ is revealed to the learner and the correct label $y_t$ is kept secret by
the adversary. Upon receiving the feature vector $x_t$, the learner is asked to
make a prediction $\widehat{y}_t$. Immediately after making its prediction, the
learner receives a feedback. In contrast with the standard full information
setting (where the feedback given is the correct label $y_t$), here the feedback
is only a binary indicator of whether the prediction was correct or not. The
protocol of the problem is formally stated below.

@@

The performance is of the learner is measured by its cumulative number of
mistakes $\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$.
The bandit feedback is a natural feedback model motivated by applications
including online advertising and online content optimization.

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}.
\chicheng{I wonder if we should introduce the two notions of linear separability here,
because strong linear separability is also worth mentioning in the full-information setting.
}
\david{The formal definitions are in a section of their own.}

### 2019-01-07 arxiv-version/main.tex
Some clean ups

\begin{abstract}
We study the problem of efficient online multiclass linear classification with bandit feedback, where all examples belong to one of $K$ classes and lie in the $d$-dimensional Euclidean space.
Previous works have left open the challenge of designing efficient algorithms with finite mistake bounds
when the data is linearly separable by a margin $\gamma$.
In this work, we take a first step towards this problem.
We identify two notions of linear separability:
one-versus-all (OvA) linear separability and weakly linear separability.
\begin{enumerate}
\item Under the OvA
linear separability condition, we design an efficient algorithm that achieves a near-optimal mistake bound of $O(\frac{K}{\gamma^2})$.
\item Under the more challenging weakly linear separability condition, we design an efficient algorithm with a mistake bound of $2^{\tilde{O}(\min(K \log^2 \frac 1 \gamma, \log K \sqrt{\frac 1 \gamma})) }$ \footnote{We use the notation $\tilde{O}(f(\cdot)) = O( f(\cdot) \polylog(f(\cdot)) )$.}.
Our algorithm is based on a reduction from weakly linear separability to OvA linear separability via an infinite dimensional monomial feature map, which in turn is inspired by ~\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}

%We focus on the setting where there are $K$ classes, and there is a $d$-dimensional multiclass linear classifier that separates all the examples shown with a positive margin $\gamma$.
%$2^{O(\min(K \log^2 K \log^2 \frac 1 \gamma, \log K \log \log K \sqrt \frac 1 \gamma \log \frac 1\gamma))}$
%$\min( (\log K \sqrt{\frac 1 \gamma})^{O(\log K \sqrt{\frac 1 \gamma})}, K^{O(K \log K \log^2 \frac1\gamma)} )$.

\end{abstract}

\input{introduction}

\input{related-work}

\input{notions-of-linear-separability}

\input{algorithm-for-strongly-linearly-separable-data}

\input{from-linear-separability-to-strong-linear-separability}

\input{proofs-of-polynomial-approximation-theorems}

\input{nearest-neighbor-algorithm}

\input{np-hardness-of-the-weak-labeling-problem}

\input{mistake-lower-bound-for-ignorant-algorithms}

>>>>

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We identify two notions of linear separability:
strong linear separability and weak linear separability.
\begin{enumerate}
\item Under the strong linear separability condition, we design an efficient
algorithm that achieves a near-optimal mistake bound of $O\left(\frac{K}{\gamma^2} \right)$.

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a reduction from weak linear separability to strong linear separability
via an infinite dimensional monomial feature map, which in turn is inspired by
\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}
\end{abstract}

\input{introduction}
\input{related-work}
\input{notions-of-linear-separability}
\input{algorithm-for-strongly-linearly-separable-data}
\input{from-linear-separability-to-strong-linear-separability}
\input{proofs-of-polynomial-approximation-theorems}
\input{nearest-neighbor-algorithm}
\input{np-hardness-of-the-weak-labeling-problem}
\input{mistake-lower-bound-for-ignorant-algorithms}

### 2019-01-07 arxiv-version/related-work.tex
Some clean ups

\section{Related work}
\label{section:related-work}
The problem of bandit multiclass linear classification is initially formulated
 by the pioneering work of~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. In that work,
 two computationally
 inefficient algorithms that work in the $\gamma$-weak linearly separable setting are proposed: one with a mistake bound of
 $O(k^2 d \ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of $\tilde{O}(\frac{k^2}{\gamma^2} \ln T)$.
 In addition, the authors propose the Banditron algorithm, a computationally efficient
 algorithm that has a $O(T^{2/3})$ regret in the general setting, and has a
 $O(\sqrt{T})$ mistake bound in the $\gamma$-weak linearly separable setting.
The polynomial dependence of Banditron's mistake bound on the time horizon is undesirable
for problems with a long time horizon.
 One key open question left by~\cite{Kakade-Shalev-Shwartz-Tewari-2008} is whether
 one can deesign computationally efficient algorithms that achieve mistake bounds that match or improve over those of inefficient algorithms.
 In this paper, we take a step towards answering this question, showing that
 efficient algorithms with mistake bounds quasipolynomial in the margin parameter can be obtained.

 %Whether one can design an efficient algorithm with a
 %finite mistake bound that has no dimensionality dependence is mentioned as an open problem in
 %~\cite{Kakade-Shalev-Shwartz-Tewari-2008}, where in this paper we provide a positive answer.
 %$O(k^2 d \ln(\frac{d}{\gamma^2}))$

Many works consider bandit multiclass classification in the general non-separable setting.
\cite{Abernethy-Rakhlin-2009} poses the open problem of whether one can design
and efficient algorithm to get a $O(\sqrt{T})$ regret against some reasonable loss functions.
\cite{CrammerG13} shows that such an algorithm can be obtained, under the assumption that the distribution of the label $y$ conditioned on feature $x$ satisfies certain parametric noise assumption.
\cite{Hazan-Kale-2011} developed the Newtron algorithm, which has a regret
between $O(\ln T)$ and $O(T^{2/3})$ against the multiclass logistic loss, where the exact order of the regret depends on the diameter of the benchmark class. In particular, if the diameter is $O(\ln T)$, its regret bound would become
$O(T^{2/3})$.
The SOBA algorithm by \cite{Beygelzimer-Orabona-Zhang-2017} achieves a regret of $\tilde{O}(\sqrt{T})$ against the $\eta$-loss~\cite{Orabona-Cesa-Bianchi-Gentile-2012}; in addition, their regret bound does not depend sensitively on the diameter of the benchmark class.
\cite{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a regret of $\tilde{O}(\sqrt{T})$ against the multiclass logistic loss, where it doubly-exponentially improves over \cite{Hazan-Kale-2011}'s regret on its dependence on the diameter of the benchmark class.
Recently, \cite{Foster-Krishnamurthy-2018} developed a rich theory on contextual bandits with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF} \frac 1 K \sum_{i=1}^K \phi( f_i(x) )$, where $\calF$ contains function $f$ such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$,
and $\phi(s) = \max(1 - \frac s \gamma, 0)$ and $\min(1, \max(1 - \frac s \gamma, 0))$.
On one hand, it gives information-theoretic mistake upper bounds under various settings of $\calF$, for instance parametric and nonparametric classes; On the other hand, it gives an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of $\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$.

\cite{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass boosting under bandit feedback, where one can view online boosting as online linear classification by treating each base hypothesis as a separate feature.
\cite{Chen-Lin-Lu-2014}'s online weak learning condition implies that there is a convex combination of base hypotheses that one-versus-all separates all examples with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$ mistake bound.
\cite{Zhang-Jung-Tewari-2018} considers an online weak learning condition that implies that there is a convex combination of base hypotheses that weakly separates all examples with a margin (See~\cite{Mukherjee-Schapire-2013}, Theorem 3). Under this condition, it gives an algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge parameter; in addition, it gives an adaptive algorithm with a $O(T^{3/4})$ mistake bound.

%\cite{Chen-Chen-Zhang-Chen-Zhang-2009} studies the approach of reducing bandit multiclass learning to online binary classification using one-versus-all reduction. They show that

%Their results do not imply a finite mistake bound in the weakly separable setting, in that the benchmark loss can still be $\Omega(T)$.

%\begin{itemize}
%\item \cite{Abernethy-Rakhlin-2009}

%\item \cite{Chen-Chen-Zhang-Chen-Zhang-2009}

%\item \cite{Hazan-Kale-2011}

%\item \cite{Beygelzimer-Orabona-Zhang-2017}

%\item \cite{Foster-Kale-Luo-Mohri-Sridharan-2018}

%\item \cite{Foster-Krishnamurthy-2018}
%\end{itemize}

%TODO

>>>>

\section{Related work}
\label{section:related-work}

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the $\gamma$-weak
linearly separable setting are proposed: one with a mistake bound of $O(k^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{k^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret in the general setting, and has a $O(\sqrt{T})$ mistake
bound in the $\gamma$-weak linearly separable setting. The polynomial dependence
of \textsc{Banditron}'s mistake bound on the time horizon is undesirable for
problems with a long time horizon. One key open question left
by~\cite{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can deesign
computationally efficient algorithms that achieve mistake bounds that match or
improve over those of inefficient algorithms. In this paper, we take a step
towards answering this question, showing that efficient algorithms with mistake
bounds quasipolynomial in the margin parameter can be obtained.

%Whether one can design an efficient algorithm with a finite mistake bound that
%has no dimensionality dependence is mentioned as an open problem in
%~\cite{Kakade-Shalev-Shwartz-Tewari-2008}, where in this paper we provide a
%positive answer. $O(k^2 d \ln(\frac{d}{\gamma^2}))$

Many works consider bandit multiclass classification in the general
non-separable setting. \cite{Abernethy-Rakhlin-2009} poses the open problem of
whether one can design and efficient algorithm to get a $O(\sqrt{T})$ regret
against some reasonable loss functions. \cite{Crammer-Gentile-2013} show that
such an algorithm can be obtained, under the assumption that the distribution of
the label $y_t$ conditioned on feature vector $x_t$ satisfies certain parametric
noise assumption. \cite{Hazan-Kale-2011} developed the \textsc{Newtron}
algorithm, which has a regret between $O(\ln T)$ and $O(T^{2/3})$ against the
multiclass logistic loss, where the exact order of the regret depends on the
diameter of the benchmark class. In particular, if the diameter is $O(\ln T)$,
its regret bound would become $O(T^{2/3})$. The SOBA algorithm by
\cite{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
$\widetilde{O}(\sqrt{T})$ against the
$\eta$-loss~\cite{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, their regret
bound does not depend sensitively on the diameter of the benchmark class.
\cite{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
it doubly-exponentially improves over \cite{Hazan-Kale-2011}'s regret on its
dependence on the diameter of the benchmark class. Recently,
\cite{Foster-Krishnamurthy-2018} developed a rich theory on contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\frac{1}{K} \sum_{i=1}^K \phi( f_i(x) )$, where $\calF$ contains function $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ and $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
for instance parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$.

\cite{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass
boosting under bandit feedback, where one can view online boosting as online
linear classification by treating each base hypothesis as a separate feature.
\cite{Chen-Lin-Lu-2014}'s online weak learning condition implies that there is a
convex combination of base hypotheses that one-versus-all separates all examples
with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
mistake bound. \cite{Zhang-Jung-Tewari-2018} considers an online weak learning
condition that implies that there is a convex combination of base hypotheses
that weakly separates all examples with a margin
(See~\cite{Mukherjee-Schapire-2013}, Theorem 3). Under this condition, it gives
an algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge
parameter. In addition, it gives an adaptive algorithm with a $O(T^{3/4})$
mistake bound.

%\cite{Chen-Chen-Zhang-Chen-Zhang-2009} studies the approach of reducing bandit
%multiclass learning to online binary classification using one-versus-all
%reduction. They show that

%Their results do not imply a finite mistake bound in the weakly separable
%setting, in that the benchmark loss can still be $\Omega(T)$.
